Let's Get Movin': Frontiers of Optimal Transport

### 2018-04-25 reports/paper/paper.tex
more on paper; genetic section

Let \[H(\b P)= -\sum_{i,j} P_{ij}(\log P_{ij} - 1)\] be the \emph{entropy} of
 the transport matrix $\b P$. Observe that \[
\nabla^2_{\b P} (-H(\b P)) - I \text{ is positive semidefinite},
\]
and so $-H(\b P)$ is 1-strongly convex. We regularize the objective of 
\eqref{eq:ot_problem} by $-\epsilon H(\b P):$ \begin{equation}
\label{eq:entropic}
  \min_{\b P \in \feasible} \sum_{i,j}P_{ij}C_{ij} - \epsilon H(\b P) 
\tag{Entropic Regularization}
\end{equation}
so as to make the objective $\epsilon$-strongly convex. \emph{Entropic
 regularization} approximates \eqref{eq:ot_problem} by solving 
\eqref{eq:entropic}, where \[
\argmin_{\b P \in \feasible} \sum_{i,j}P_{ij}C_{ij} - \epsilon H(\b P) 
 {\underset{n\to\infty}{\longrightarrow}}
 \argmin_{\b P \in \feasible} \sum_{i,j}P_{ij}C_{ij},
\]
where the convergence is towards the maximal-entropy solution of 
\eqref{eq:ot_problem}.

@@

@@

\subsection{Genetic Algorithms (Hokey)} % (fold)
\label{sub:genetic_algorithms_}

>>>>

Let \[H(\b P)= -\sum_{i,j} P_{ij}(\log P_{ij} - 1)\] be the \emph{entropy} of
 the transport matrix $\b P$. Observe that \[
\nabla^2_{\b P} (-H(\b P)) - I \text{ is positive semidefinite},
\]
and so $-H(\b P)$ is 1-strongly convex. We regularize the objective of 
\eqref{eq:ot_problem} by $-\epsilon H(\b P):$ \begin{equation}
\label{eq:entropic}
  \min_{\b P \in \feasible} \sum_{i,j}P_{ij}C_{ij} - \epsilon H(\b P) 
\tag{Entropic Regularization}
\end{equation}
so as to make the objective $\epsilon$-strongly convex. \emph{Entropic
 regularization} approximates \eqref{eq:ot_problem} by solving 
\eqref{eq:entropic}, where \[
\argmin_{\b P \in \feasible} \sum_{i,j}P_{ij}C_{ij} - \epsilon H(\b P) 
 {\underset{n\to\infty}{\longrightarrow}}
 \argmin_{\b P \in \feasible} \sum_{i,j}P_{ij}C_{ij},
\]
where the convergence is towards the maximal-entropy solution of 
\eqref{eq:ot_problem}. The convergence property makes \eqref{eq:entropic} an
 attractive alternative to solving optimal transport; other reasons why 
\eqref{eq:entropic} is useful is explored in a seminal article by 
\cite{cuturi2013sinkhorn}.

@@

\begin{rmk}
  We do not need to limit ourselves to using coordinate descent to minimize $f$.
 For instance
\cite{brauer2017sinkhorn} uses Newton's method, and achieves similar convergence
 performance as \cref{alg:sinkhorn}.
\end{rmk}

@@

\subsection{Genetic Algorithm} % (fold)
\label{sub:genetic_algorithms_}

Recent advances in machine learning points towards the versatility of genetic
 algorithms, even as a competitive alternative to deep learning based on
 gradient methods \cite{such2017deep}. Inspired by these advances, we consider
 the following genetic algorithm

\begin{algorithm}
\caption{A toy genetic algorithm for optimal transport}
\label{alg:genetic}

  \begin{algorithmic}
\Function{GeneticOT}{$\phi, \gamma, k, \b a, \b b, T$}
\State $\mathsf{initial} \gets \b a \b b^T$
\State $\mathsf{family} \gets \mathsf{Array}[\phi]$, initialized to $
\mathsf{initial}$
\State $\mathsf{children} \gets \mathsf{Array}[\phi\gamma]$, initialized to
 empty
  \For{$\mathsf{generation}$ in $1:T$}
    \For{$\mathsf{person}$ in $\mathsf{family, \gamma}$}
      \For{$\mathsf{person}'$ in \Call{Subset}{$\mathsf{family}$}} \Comment
{\textsc{Subset} chooses a subset of size $\gamma$}
\State $\mathsf{child} \gets$ \Call{Marry}{$\mathsf{person}, 
\mathsf{person}'$}
\State $\mathsf{child} \gets$ \Call{Perturb}{$\mathsf{child},k$}
\State Add $\mathsf{child}$ to $\mathsf{children}$
\EndFor
\EndFor
\State Choose the $\phi$ elements in $\mathsf{children}$ with minimal cost for
 \eqref{eq:ot_problem} and set to $\mathsf{family}$
\EndFor
\State\Return{element in $\mathsf{family}$ that achieves minimal cost}
\EndFunction
  \end{algorithmic}
\end{algorithm}
Here \textsc{Marry} picks two probability matrices and averages them
 elementwise, and \textsc{Perturb} chooses random indices $i_1,i_2, j_1,j_2$ and
 moves mass in the probability matrix in such a way that preserves the marginal
 distributions and respects the constraint that $0\le p \le 1$ for all
 components $p$. Such perturbations are repeated $k$ times. $\phi$ is a
 parameter that controls the size of families, and $\gamma$ is a parameter that
 decides the number of offsprings. At each generation, we take the top $\phi$
 matrices that achieve the lowest cost. We hoped that the population drifts
 towards the optimal transport matrix as the number of generation increases.

The empirical results we obtain are less than encouraging, however. We plot the
 optimal cost each generation with $n = m = 4, \phi=100, \gamma =5$ for various
 values of $k$ in \cref{fig:genetic_fail}. Curiously, the improvement in the
 first generation is always large, with a magnitude of improvement increasing
 in $k$, but the algorithm seems to achieve some non-optimal steady state as
 generation goes on. Of course, \cref{alg:genetic} can potentially be improved
 by thinking about the design of \textsc{Marry} and \textsc{Perturb} more
 carefully, perhaps leveraging some information about $\b C$.

\begin{figure}[tb]
  \centering
  \includegraphics{figs/genetic.eps}
  \caption{Non-convergence of \cref{alg:genetic}}
  \label{fig:genetic_fail}
\end{figure}
